---
title: "Course Project 1"
author: "Manuel Chacón"
date: "28 de mayo de 2019"
output:
 html_document:
  keep_md: true
---
```{r}
library(dplyr)
library(ggplot2)
```




## Loading and preprocessing the data
```{r}
data <- read.csv('activity.csv')
```



## What is mean total number of steps taken per day?

we create  data_step, It's the number steps per day.
```{r}
data_step<-data %>% 
          group_by(date) %>% 
          summarize(steps_day=sum(steps), steps_mean = mean(steps), steps_median = median(steps,na.rm = TRUE))

```

```{r}
ggplot(data_step, aes(steps_day))+
  geom_histogram(color = 'red')+ ggtitle('Steps per days')+ylab('Frecuency')+xlab('Number of steps')
```

```{r}
data_mean <- data_step %>% 
  select(date, steps_mean, steps_median)
```

## What is the average daily activity pattern?
```{r}
data_interval<-data %>% 
  group_by(interval) %>% 
  summarize(interval_mean = mean(steps, na.rm = TRUE))
```


```{r}
ggplot(data_interval, aes(interval, interval_mean))+geom_line(color='blue')+
  ggtitle('Average per interval')+xlab('Interval')+ylab('Average per interval')
```
```{r}
data_interval %>% 
  select(interval, interval_mean) %>% 
  filter(interval_mean == max(interval_mean))
```

## Imputing missing values
1- We calculate total number of missing value
```{r}
data %>%
  summarize(steps_na=sum(is.na(steps)),
            date_na=sum(is.na(date)), 
            interval_na=sum(is.na(interval)))
```

2.-We create a new data that is equal to the original but with the missing data filled in.
```{r}
data_copy<-data

data_copy$steps[is.na(data_copy$steps)] <- mean(na.omit(data_copy$steps))
```
3.-Make a histogrom of the total number of steps taken each day and calculate and report the mean and median total number of steps taken per day.
```{r}
par(mfrow=c(1,2))
data %>%
  group_by(date) %>% 
  summarize(total_steps = sum(steps)) %>% 
  ggplot(aes(total_steps))+
  geom_histogram(color = 'red') +
  ggtitle('With NA')+
  xlab('Number of steps')+
  ylab('Frecuency')

data_copy %>%
  group_by(date) %>% 
  summarize(total_steps = sum(steps)) %>% 
  ggplot(aes(total_steps))+
  geom_histogram(color = 'blue')+
  ggtitle('Without NA')+
  xlab('Number of steps')+
  ylab('Frecuency')
```
```{r warning=FALSE}
data_copy %>% 
  group_by(date) %>% 
  summarize(steps_day=sum(steps), steps_mean = mean(steps), steps_median = median(steps))
```



## Are there differences in activity patterns between weekdays and weekends?

1.-Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
data_week<-data_copy %>% 
  mutate(Weekday = ifelse(weekdays(as.Date(date)) %in% c("lunes","martes",
                                                         'miercoles','jueves','viernes')
                          ,"weekday",'weekend' ) )
```

2.-Make a panel plot containing a time series plot of the 5-minute interval(x-axis) and the average of steps taken, averaged across all weekday days or weekend days(y-axis).

```{r}
data_weekday <-data_week %>% 
  filter(Weekday == 'weekday') %>% 
  group_by(interval) %>% 
  summarize(interval_step = mean(steps))

data_weekend <-data_week %>% 
  filter(Weekday == 'weekend') %>% 
  group_by(interval) %>% 
  summarize(interval_step = mean(steps))


par(mfrow=c(1,2))
ggplot(data_weekday, aes(interval, interval_step))+geom_line()+ggtitle('Average per interval on weekday')+xlab('Interval')+ylab('Average per interval')

ggplot(data_weekend, aes(interval, interval_step))+geom_line()+ggtitle('Average per interval on weekend')+xlab('Interval')+ylab('Average per interval')
```

